---
title: "Red Wine Quality Exploratory Data Analysis"
author: "by Paul Smith"
output: html_document
---
# Contents:
1. [Introduction](#introduction)
2. [Univariate Plots Section](#univariate_plots_section)
3. [Univariate Analysis](#univariate_analysis)
4. [Bivariate Plots Section](#bivariate_plots_section)
5. [Bivariate Analysis](#bivariate_analysis)
6. [Multivariate Plots Section](#multivariate_plots_section)
7. [Multivariate Analysis](#multivariate_analysis)
8. [Final Plots and Summary](#final_plots_and_summary)
9. [Reflection](#reflection)

```{r echo=FALSE, message=FALSE, warning=FALSE, required_packages}
# Load all of the required packages in this code chunk.
library(ggplot2)
library(knitr)
library(dplyr)
library(corrplot)
```

```{r echo=FALSE, message=FALSE, warning=FALSE, load_dataset}
# Load the provided dataset.
red_wine_ds <- read.csv('~/wgu-c751/wineQualityReds.csv', row.names = 1)
```

------

<a name="introduction"></a>
**1. Introduction**

This report explores a dataset containing quality ratings and chemical
properties for 1,599 red wines (all from the Portuguese *Vinho Verde* red wine
variant). The quality ratings were derived from the median of at least 3 wine
experts who rated the quality of each wine between 0 (very bad) and 10 (very excellent).

The dataset contains 12 variables - 11 input numerical variables based on
physiochemical tests, and 1 categorical output variable (quality) based on
sensory data:

 1. *fixed acidity* - most acids involved with wine, or fixed, or nonvolatile.
 2. *volatile acidity* - the amount of acetic acid in wine.
 3. *citric acid* - found in small quantities: adds 'freshness' and flavor.
 4. *residual sugar* - the amount of sugar remaining after fermentation stops.
 5. *chlorides* - the amount of salt in the wine.
 6. *free sulfur dioxide* - free form of sulfur dioxide.
 7. *total sulfur dioxide* - amount of free and bound sulfur dioxide.
 8. *density* - the density of the wine.
 9. *pH* - how acidic or basic a wine is.
10. *sulphates* - an additive that contributes to sulfur dioxide.
11. *alcohol* - the alcohol percentage of the wine.
12. *quality* - the output variable based on sensory data (between 0 and 10).

------

<a name="univariate_plots_section"></a>
**2. Univariate Plots Section**

The dataset will be explored using a single variable at a time. The goal is to
find out which property (or properties) affects the quality of the wine. The
analysis will start with the quality variable, followed by an analysis of each
of the input variables.

*Wine Quality:*

Based on a visual inspection,the quality values of 5 and 6 are the most common:

```{r echo=FALSE, message=FALSE, warning=FALSE, wine_quality}
# Wine quality.
ggplot(aes(x = quality, color = I('white')), data = red_wine_ds)+
  geom_bar()+ scale_x_continuous(breaks = seq(3, 8, 1))+
  xlab("Wine Quality")+
  ylab("Count")
```

We will confirm that the visual inspection is correct:

```{r echo=FALSE, message=FALSE, warning=FALSE, most_quality_values}
# Proportion of wines with a quality value of 5 or 6.
sum(red_wine_ds$quality == 5 | red_wine_ds$quality == 6) /
  nrow(red_wine_ds) * 100
```

82.49% (rounded to two decimal places) of the wines are of quality value 5 and
6.

*Fixed Acidity:*

The graph peaks at approximately 7.5:

```{r echo=FALSE, message=FALSE, warning=FALSE, fixed_acidity}
# Fixed acidity.
ggplot(aes(x = fixed.acidity, color = I('white')), data = red_wine_ds)+
  geom_histogram(bins = 35)+
  xlab("Fixed Acidity")+
  ylab("Count")+
  scale_x_continuous(breaks = seq(4.5, 16, 1))
```

Calculate the proportion of fixed acidity that lies in the 7 to 8 range:

```{r echo=FALSE, message=FALSE, warning=FALSE, range_7_8}
# Proportion of fixed acidity lies in the 7 to 8 range.
sum(red_wine_ds$fixed.acidity >= 7 & red_wine_ds$fixed.acidity <= 8)/
  nrow(red_wine_ds) * 100
```

Approximately 34.15% (rounded to two decimal places) of fixed acidity lies in
the 7 to 8 range.

*Volatile Acidity:*

Very few wines have a volatile acidity of more than 1:

```{r echo=FALSE, message=FALSE, warning=FALSE, volatile_acidity}
# Volatile acidity.
ggplot(aes(x = volatile.acidity, color = I('white')), data = red_wine_ds)+
  geom_histogram(bins = 35)+
  xlab("Volatile Acidity")+
  ylab("Count")+
  scale_x_continuous(breaks = seq(0, 1.7, 0.1))
```

These are likely outliers. After removing these outliers we see:

```{r echo=FALSE, warning=FALSE, message=FALSE, volatile_acidity_no_outliers}
# Volatile acidity without outliers.
ggplot(aes(x = volatile.acidity, color = I('white')), data = red_wine_ds)+
  geom_histogram(bins = 35)+
  xlab("Volatile Acidity (No Outliers)")+
  ylab("Count")+
  scale_x_continuous(breaks = seq(0, 1.1, 0.1), limits = c(0.1, 1))
```

Most of the volatile acidity is from 0.3 to 0.7 - a normal distribution with
some peaks.

*Citric Acid:*

Peaks appear at 0 and 0.48:

```{r echo=FALSE, message=FALSE, warning=FALSE, citric_acid}
# Citric acid.
ggplot(aes(x = citric.acid, color = I('white')), data = red_wine_ds)+
  geom_histogram(binwidth = 0.01)+
  xlab("Citric Acid")+
  ylab("Count")+
  scale_x_continuous(breaks = seq(0, 1, 0.05))
```

By far the biggest peak is value 0.  Calculate how many wines have a value of 0:

```{r echo=FALSE, message=FALSE, warning=FALSE, wines_with_value_0}
# Number of wines with value of 0.
sum(red_wine_ds$citric.acid == 0)
```

132 wines have a citric acid value of 0.  Calculate what proportion of wines
have a value below 0.5:

```{r echo=FALSE, message=FALSE, warning=FALSE, value_below_point_5}
# Proportion of wines with a value below 0.5.
sum(red_wine_ds$citric.acid > 0.0 & red_wine_ds$citric.acid < 0.5) /
  nrow(red_wine_ds) * 100
```

78.61% (rounded to two decimal places) of the wines have citric acid values
below 0.5 (including 132 wines with a value of 0 as calculated earlier).

*Residual Sugar:*

Outliers are observed:

```{r echo=FALSE, message=FALSE, warning=FALSE, residual_sugar}
# Residual sugar.
ggplot(aes(x = residual.sugar, color = I('white')), data = red_wine_ds )+
  geom_histogram(bins = 35)+
  xlab("Residual Sugar")+
  ylab("Count")+
  scale_x_continuous(breaks = seq(0, 16, 1))
```

After removing the outliers we see:

```{r echo=FALSE, message=FALSE, warning=FALSE, residual_sugar_no_outliers}
# Residual sugar without outliers.
ggplot(aes(x = residual.sugar, color = I('white')),data = red_wine_ds )+
  geom_histogram(bins = 30)+
  xlab("Residual Sugar (No Outliers)")+
  ylab("Count")+
  scale_x_continuous(limits = c(1, 9), breaks = seq(1, 9, 0.5))
```

The values peak at approximately 2 for residual sugar.

*Chlorides:*

Outliers are observed:

```{r echo=FALSE, message=FALSE, warning=FALSE, chlorides}
# Chlorides.
ggplot(aes(x = chlorides, color = I('white')), data = red_wine_ds) + geom_histogram(bins = 50) +
  xlab("Chlorides")+
  ylab("Count")+
  scale_x_continuous(breaks = seq(0, 0.7, 0.1))
```

Removing the outliers we see:

```{r echo=FALSE, message=FALSE, warning=FALSE, chlorides_no_outliers}
# Chlorides without outliers.
ggplot(aes(x = chlorides, color = I('white')), data = red_wine_ds) +
  xlab("Chlorides (No Outliers)")+
  ylab("Count")+
  geom_histogram(bins = 30)+
  scale_x_continuous(breaks = seq(0.01, 0.25, 0.01), limits= c(0.03, 0.25))
```

The chlorides peak at approximately 0.75.

*Free Sulfur Dioxide:*

Outliers are observed:

```{r echo=FALSE, message=FALSE, warning=FALSE, free_sulfur_dioxide_outliers}
# Free sulphur dioxide.
ggplot(aes(x = free.sulfur.dioxide, color = I('white')), data = red_wine_ds) +
  geom_histogram(bins = 35) +
  xlab("Free Sulfur Dioxide")+
  ylab("Count")+
  scale_x_continuous(breaks = seq(0, 80, 5))
```

Removing the outliers shows:

```{r echo=FALSE, message=FALSE, warning=FALSE, free_sulfur_dioxide}
# Free sulphur dioxide without outliers.
ggplot(aes(x = free.sulfur.dioxide, color = I('white')), data = red_wine_ds)+
  geom_histogram(binwidth = 1)+
  xlab("Free Sulfur Dioxide (No Outliers)")+
  ylab("Count")+
  scale_x_continuous(limits = c(2,55),breaks = seq(0,55,2))
```

The sulfur dioxide peaks at approximately 6.

*Total Sulfur Dioxide:*

Outliers are observed:

```{r echo=FALSE, message=FALSE, warning=FALSE, total_sulfur_dixoide_outliers}
# Total sulfur dioxide.
ggplot(aes(x = total.sulfur.dioxide, color = I('white')), data = red_wine_ds) +
  xlab("Total Sulfur Dioxide")+
  ylab("Count")+
  geom_histogram(binwidth = 5)
```

Removing the outliers shows:

```{r echo=FALSE, message=FALSE, warning=FALSE, total_sulfur_dixoide}
# Total sulfur dioxide wihout outliers.
ggplot(aes(x = total.sulfur.dioxide, color=I('white')), data=red_wine_ds)+
  geom_histogram(binwidth = 3)+
  xlab("Total Sulfur Dioxide (No Outliers)")+
  ylab("Count")+
  scale_x_continuous(limits = c(0,175))
```

The total sulfur dioxide peaks between 15 and 25.

*Density:*

Density has minimal variability:

```{r echo=FALSE, message=FALSE, warning=FALSE, density}
# Density.
ggplot(aes(x = density, color = I('white')), data = red_wine_ds)+
  geom_histogram(bins = 35)+
  xlab("Density")+
  ylab("Count")+
  scale_x_continuous(breaks = seq(0.09, 1.1, 0.001))
```

Calculate the proportion of wines with density between 0.9945 and 0.9985:

```{r echo=FALSE, message=FALSE, warning=FALSE, show_minimal_variability}
# Minimal variability.
sum(red_wine_ds$density >= 0.9945 & red_wine_ds$density <= 0.9985)/
  nrow(red_wine_ds) * 100
```

74.61% (rounded to two decimal places) of the wines have a density between
0.9945 and 0.9985.

*pH:*

Outliers are observed:

```{r echo=FALSE, message=FALSE, warning=FALSE, ph}
# pH.
ggplot(aes(x = pH, color = I('white')), data = red_wine_ds)+
  geom_histogram(bins = 35)+
  xlab("pH")+
  ylab("Count")+
  scale_x_continuous(breaks = seq(3, 4, 0.1))
```

Without outliers we see:

```{r echo=FALSE, message=FALSE, warning=FALSE, ph_no_outliers}
# pH without outliers.
ggplot(aes(x = pH, color = I('white')), data = red_wine_ds)+
  geom_histogram(bins = 35)+
  xlab("pH (No Outliers)")+
  ylab("Count")+
  scale_x_continuous(limits = c(2.8,3.8))
```

Calculate whether the majority of wines lie within the pH range of 3.2 to 3.4:

```{r echo=FALSE, message=FALSE, warning=FALSE, common_ph}
# Most wines pH value.
sum(red_wine_ds$pH >= 3.2 & red_wine_ds$pH <= 3.4) /
  nrow(red_wine_ds) * 100
```

53.97% of wines have a pH of approximately 3.2 to 3.4.

*Sulphates:*

Outliers are observed:

```{r echo=FALSE, message=FALSE, warning=FALSE, sulphates}
# Sulphates.
ggplot(aes(x = sulphates, color = I('white')), data = red_wine_ds)+geom_histogram(bins = 30)+
  xlab("Sulphates")+
  ylab("Count")
```

Once removed we see:

```{r echo=FALSE, message=FALSE, warning=FALSE, sulphates_no_outliers}
# Sulphates without outliers.
ggplot(aes(x = sulphates, color = I('white')), data = red_wine_ds)+
  geom_histogram(binwidth = 0.01)+
  xlab("Sulphates (No Outliers)")+
  ylab("Count")+
  scale_x_continuous(limits = c(0.4, 0.9))
```

Sulphate values peak at approximately 0.6.

*Alcohol:*

Outliers were observed:

```{r echo=FALSE, message=FALSE, warning=FALSE, alcohol}
# Alcohol.
ggplot(aes(x = alcohol, color = I('white')), data = red_wine_ds)+
  geom_histogram(bins=35)+
  xlab("Alcohol")+
  ylab("Count")
```

Once removed we see:

```{r echo=FALSE, message=FALSE, warning=FALSE, alcohol_no_outliers}
# Alcohol without outliers.
ggplot(aes(x = alcohol, color = I('white')), data = red_wine_ds)+
  geom_histogram(binwidth = 0.1)+
  xlab("Alcohol (No Outliers)")+
  ylab("Count")+
  scale_x_continuous(limits = c(9, 13), breaks = seq(9, 13, 0.3))
```

There is a peak at approximately 9.5.

------

<a name="univariate_analysis"></a>
**3. Univariate Analysis**

**What is the structure of your dataset?**

The red wine dataset contains 1,599 observations and 12 variables.  11 of them
are input variables based on physiochemical tests.  The remaining variable,
quality, is the output variable based on sensory data. Each observation
corresponds to one particular wine.

Here is the dataset structure:

```{r echo=FALSE, message=FALSE, warning=FALSE, dataset_structure}
# Structure of the dataset.
str(red_wine_ds)
```

**What is/are the main feature(s) of interest in your dataset?**

The main feature of interest is that 82.49% of the wines are of quality value 5
and 6. I want to find out which of the input variables lead to a high quality
wine. Possible variables that could influence the quality of wines are:

1. 34.15% of the fixed acidity values lie between 7 and 8.
2. 78.61% of the citric acid values lie below 0.5.
3. Sulphates peak at approximately 0.6.
4. Alcohol peaks at approximately 9.5.

**What other features in the dataset do you think will help support your
investigation into your feature(s) of interest?**

The 4 input variables listed previously could contribute to wine quality.
The bivariate analysis will look at how each of the attributes is distributed
with a given quality value.

**Did you create any new variables from existing variables in the dataset?**

I did not create any new variables.

**Of the features you investigated, were there any unusual distributions?
Did you perform any operations on the data to tidy, adjust, or change the form
of the data? If so, why did you do this?**

Once the outliers were removed from the chlorides and residual sugar graphs,
what appeared to be left-skewed graphs were in fact normal graphs.  Limits were
used to scale down the x-axis variable, as this helped to analyze the data.

------

<a name="bivariate_plots_section"></a>
**4. Bivariate Plots Section**

The univariate plots did not indicate which variables influenced wine quality.
A new strategy was implemented - a correlation graph of the variables was
plotted to determine which variables should be plotted against wine quality:

```{r echo=FALSE, message=FALSE, warning=FALSE, correlation_graph}
# Correlation graph.
M <- cor(red_wine_ds)
M2 = M
diag(M2) = NA
colnames(M2) = rep(c('Fixed Acidity',
                     'Volatile Acidity',
                     'Citric Acid',
                     'Residual Sugar',
                     'Chlorides',
                     'Free Sulfur Dioxide',
                     'Total Sulfur Dioxide',
                     'Density',
                     'pH',
                     'Sulphates',
                     'Alcohol',
                     'Quality'),
                   c(1, 1, 1, 1, 1, 1, 1, 1, 1, 1, 1, 1))
rownames(M2) = rep(c('Fixed Acidity',
                     'Volatile Acidity',
                     'Citric Acid',
                     'Residual Sugar',
                     'Chlorides',
                     'Free Sulfur Dioxide',
                     'Total Sulfur Dioxide',
                     'Density',
                     'pH',
                     'Sulphates',
                     'Alcohol',
                     'Quality'),
                   c(1, 1, 1, 1, 1, 1, 1, 1, 1, 1, 1, 1))
corrplot(M2,
         addCoef.col = 'black',
         number.cex=0.5,
         na.label = 'X',
         method =  'circle')
```

We see that volatile acidity (-0.39), citric acid (0.23), sulphates (0.25), and
alcohol (0.48) are correlated with wine quality.  These variables will be
plotted on the y-axis and wine quality will be plotted on the x-axis of a
box plot.

*Volatile Acidity vs. Wine Quality:*

```{r echo=FALSE, message=FALSE, warning=FALSE, volatile_acidity_vs_quality}
# Volatile Acidity vs. Wine Quality.
ggplot( aes(x = as.factor(quality), y = red_wine_ds$volatile.acidity),
  data = red_wine_ds)+
  geom_boxplot()+
  xlab("Wine Quality")+
  ylab("Volatile Acidity")+
  scale_x_discrete(breaks = seq(1, 10, 1))+
  scale_y_continuous(breaks = seq(0.1, 1.7, 0.1))
```

Low volatile acidity is a sign of a good quality wine. This is negatively
correlated with wine quality.

*Citric Acid vs Wine Quality:*

```{r echo=FALSE, message=FALSE, warning=FALSE, citric_acid_vs_quality}
# Citric Acid vs. Wine Quality.
ggplot( aes(x = as.factor(quality), y = red_wine_ds$citric.acid),
  data = red_wine_ds)+
  geom_boxplot()+
  xlab("Wine Quality")+
  ylab("Citric Acid")+
  scale_x_discrete(breaks = seq(1, 10, 1))+
  scale_y_continuous(breaks = seq(0, 1, 0.1))
```

Good quality wines have high levels of citric acid - this is directly
proportional to wine quality.

*Sulphates vs. Wine Quality:*

```{r echo=FALSE, message=FALSE, warning=FALSE, sulphates_vs_quality}
# Sulphates vs. Wine Quality.
ggplot(aes(x = as.factor(quality), y = red_wine_ds$sulphates),
  data = red_wine_ds)+
  geom_boxplot()+
  xlab("Wine Quality")+
  ylab("Sulphates")+
  scale_x_discrete(breaks = seq(1, 10, 1))+
  scale_y_continuous(breaks = seq(0.2, 2, 0.05), limits = c(0.3, 0.9))
```

Good quality wines have high values of sulphates - these are directly
proportional to wine quality.

*Alcohol vs. Wine Quality:*

```{r echo=FALSE, message=FALSE, warning=FALSE, alcohol_vs_quality}
# Alcohol vs. Wine Quality.
ggplot( aes( x = as.factor(quality), y = red_wine_ds$alcohol),
  data = red_wine_ds)+
  geom_boxplot()+
  xlab("Wine Quality")+
  ylab("Alcohol")+
  scale_x_discrete(breaks = seq(1, 10, 1))+
  scale_y_continuous(breaks = seq(8, 15, 0.5))
```

Good quality wines have high levels of alcohol.

------

<a name="bivariate_analysis"></a>
**5. Bivariate Analysis**

**Talk about some of the relationships you observed in this part of the
investigation. How did the feature(s) of interest vary with other features in
the dataset?**

The correlation plot brought into focus the variables (volatile acidity, citric
acid, sulphates, and alcohol) that are correlated with wine quality.  It showed
that other variables (such as chlorides or density) are not correlated.

The features of interest changed after this analysis. Fixed acidity was shown
to be not correlated.

**Did you observe any interesting relationships between the other features
(not the main feature(s) of interest)?**

An interesting relationship that was observed was that fixed acidity is
correlated with volatile acidity, citric acid, density, and pH value.  This will
be investigated further to see what the relationship is (if any) between this correlation and wine quality.

**What was the strongest relationship you found?**

Alcohol is the strongest correlation associated with wine quality. The bivariate analysis shows that good quality wines have high levels of alcohol.

------

<a name="multivariate_plots_section"></a>
**6. Multivariate Plots Section**

The correlation graph shows that fixed acidity is correlated with volatile
acidity (-0.26), citric acid (0.67), density (0.67), and pH (-0.68).  A
scatter plot of fixed acidity with the former metrics will be plotted to try and
understand the relation of these variables with wine quality.

```{r echo=FALSE, message=FALSE, warning=FALSE, reusable_function}
# Scatterplot function.
scatterplot <- function(x, y, color, xlabel, ylabel, dot_color, brewer_name)
{
  ggplot(aes(x = x, y = y, color = color), data = red_wine_ds)+
  xlab(xlabel)+
  ylab(ylabel)+
  geom_point()+
  geom_smooth(method = "lm", se = FALSE)+
  scale_color_brewer(brewer_name, palette = dot_color)+
  theme(panel.background = element_rect(fill = '#666666', colour = 'black'),
  panel.grid.major = element_line(colour = "black"),
  panel.grid.minor = element_line(colour = "black"),
  legend.key = element_rect(fill = '#666666'))
}
```

```{r echo=FALSE, message=FALSE, warning=FALSE, scatterplots_of_metrics}
# Volatile Acidity vs. Fixed Acidity vs. Quality.
scatterplot(red_wine_ds$volatile.acidity, red_wine_ds$fixed.acidity,
  as.factor(red_wine_ds$quality), 'Volatile Acidity',
  'Fixed Acidity', 'Blues', 'Wine Quality')
# Citric Acid vs. Fixed Acidity vs. Quality.
scatterplot(red_wine_ds$citric.acid, red_wine_ds$fixed.acidity,
  as.factor(red_wine_ds$quality), 'Citric Acid',
  'Fixed Acidity', 'Blues', 'Wine Quality')
# Density vs. Fixed Acidity vs. Quality.
scatterplot(red_wine_ds$density, red_wine_ds$fixed.acidity,
  as.factor(red_wine_ds$quality), 'Density',
  'Fixed Acidity', 'Blues', 'Wine Quality')
# pH vs. Fixed Acidity vs. Quality.
scatterplot(red_wine_ds$pH, red_wine_ds$fixed.acidity,
  as.factor(red_wine_ds$quality), 'pH',
  'Fixed Acidity', 'Blues', 'Wine Quality')
```

Straight line behavior shows the correlation between the x and y axes.  However,
no relationship with wine quality is apparent.

The continuous variables need to be converted into different ranges. Each
variable will be sliced into five partitions.  These partitions will be plotted
with variables that have correlated with wine quality.

```{r echo=FALSE, message=FALSE, warning=FALSE, data_partitions}
# Partition Volatile Acidity.
red_wine_ds$volatile.acidity.partition <-
  cut(red_wine_ds$volatile.acidity,
      breaks = c(0.12, 0.39, 0.52, 0.53, 0.64, 1.58))
# Partition Citric Acid.
red_wine_ds$citric.acid.partition <-
  cut(red_wine_ds$citric.acid,
      breaks = c(0, 0.09, 0.26, 0.271, 0.42, 1))
# Partition Sulphates.
red_wine_ds$sulphates.partition <-
  cut(red_wine_ds$sulphates,
      breaks = c(0.33, 0.55, 0.62, 0.6581, 0.73, 2))
# Partition Alcohol.
red_wine_ds$alcohol.partition <-
  cut(red_wine_ds$alcohol,
      breaks = c(8.4, 9.5, 10.2, 10.42, 11.1, 14.9))
```

*Alcohol and Wine Quality with other properties:*

```{r echo=FALSE, message=FALSE, warning=FALSE, alcohol_vs_quality_vs_other}
# Quality vs. Alcohol vs. sulphates.partition.
scatterplot(red_wine_ds$quality, red_wine_ds$alcohol,
  red_wine_ds$sulphates.partition, 'Wine Quality', 'Alcohol %', 'Greens',
  'Sulphates Partition')+ geom_jitter()
# Quality vs. Alcohol vs. citric.acid.partition.
scatterplot(red_wine_ds$quality, red_wine_ds$alcohol,
  red_wine_ds$citric.acid.partition, 'Wine Quality', 'Alcohol %', 'Greens',
  'Citric Acid Partition')+ geom_jitter()
# Quality vs. Alcohol vs. volatile.acidity.partition.
scatterplot(red_wine_ds$quality, red_wine_ds$alcohol,
  red_wine_ds$volatile.acidity.partition, 'Wine Quality', 'Alcohol %', 'Greens',
  'Volatile Acidity Partition')+ geom_jitter()
```

These plots show that high sulphates and high alcohol influence wine quality, as
does high citric acid and high alcohol.

*Volatile Acidity and Wine Quality with other properties:*

```{r echo=FALSE, message=FALSE, warning=FALSE, v_acidity_vs_quality_vs_other}
# Quality vs. Volatile Acidity vs. alcohol.partition.
scatterplot(red_wine_ds$quality, red_wine_ds$volatile.acidity,
  red_wine_ds$alcohol.partition, 'Wine Quality', 'Volatile Acidity', 'Reds',
  'Alcohol Partition')+ geom_jitter()
# Quality vs. Volatile Acidity vs. citric.acid.partition.
scatterplot(red_wine_ds$quality, red_wine_ds$volatile.acidity,
  red_wine_ds$citric.acid.partition, 'Wine Quality', 'Volatile Acidity', 'Reds',
  'Citric Acid Partition')+ geom_jitter()
# Quality vs. Volatile Acidity vs. sulphates.partition.
scatterplot(red_wine_ds$quality, red_wine_ds$volatile.acidity,
  red_wine_ds$sulphates.partition, 'Wine Quality', 'Volatile Acidity', 'Reds',
  'Sulphates Partition')+ geom_jitter()
```

These plots show that lower values of volatile acidity influence wine quality.

*Citric Acid and Wine Quality with other properties:*

```{r echo=FALSE, message=FALSE, warning=FALSE, citric_acid_vs_quality_vs_other}
# Quality vs. Citric Acid vs. alcohol.partition.
scatterplot(red_wine_ds$quality, red_wine_ds$citric.acid,
  red_wine_ds$alcohol.partition, 'Wine Quality', 'Citric Acid', 'BuPu',
  'Alcohol Partition')+ geom_jitter()
# Quality vs. Citric Acid vs. volatile.acidity.partition.
scatterplot(red_wine_ds$quality, red_wine_ds$citric.acid,
  red_wine_ds$volatile.acidity.partition, 'Wine Quality', 'Citric Acid', 'BuPu',
  'Volatile Acidity Partition')+ geom_jitter()
# Quality vs. Citric Acid vs. sulphates.partition.
scatterplot(red_wine_ds$quality, red_wine_ds$citric.acid,
  red_wine_ds$sulphates.partition, 'Wine Quality', 'Citric Acid', 'BuPu',
  'Sulphates Partition')+ geom_jitter()
```

These plots show that citric acid does not influence wine quality.

*Sulphates and Wine Quality with other properties:*

```{r echo=FALSE, message=FALSE, warning=FALSE, sulphates_vs_quality_vs_other}
# Quality vs. Sulphates vs. alchohol.partition.
scatterplot(red_wine_ds$quality, red_wine_ds$sulphates,
  red_wine_ds$alcohol.partition, 'Wine Quality', 'Sulphates', 'Purples',
  'Alcohol Partition')+ geom_jitter()
# Quality vs. Sulphates vs. volatile.acidity.partition.
scatterplot(red_wine_ds$quality, red_wine_ds$sulphates,
  red_wine_ds$volatile.acidity.partition, 'Wine Quality', 'Sulphates', 'Purples',
  'Volatile Acidity Partition')+ geom_jitter()
# Quality vs. Sulphates vs. citric.acid.partition.
scatterplot(red_wine_ds$quality, red_wine_ds$sulphates,
  red_wine_ds$citric.acid.partition, 'Wine Quality', 'Sulphates', 'Purples',
  'Citric Acid Partition')+ geom_jitter()
```

These plots show that high values of sulphates influence wine quality.

------

<a name="multivariate_analysis"></a>
**7. Multivariate Analysis**

**Talk about some of the relationships you observed in this part of the
investigation. Were there features that strengthened each other in terms of
looking at your feature(s) of interest?**

High alcohol contributes to good wine quality.  Adding sulphates or citric acid
will influence wine quality in a positive manner.  Wine quality is also
influenced with a lower value of volatile acidity.

**Were there any interesting or surprising interactions between features?**

Citric acid was shown to influence wine quality based on observing the
correlation plot.  However, from the multivariate plots we see that citric
acid does not influence wine quality on its own - it must also be paired with
high alcohol values to influence (positively) wine quality.

**OPTIONAL: Did you create any models with your dataset? Discuss the
strengths and limitations of your model.**

I did not create any models with the dataset.

------

<a name="final_plots_and_summary"></a>
**8. Final Plots and Summary**

## Plot One:

```{r echo=FALSE, message=FALSE, warning=FALSE, plot_one}
# Quality of wine.
ggplot( aes(quality, fill = I('Blue')), data = red_wine_ds)+
  geom_bar()+
  xlab("Wine Quality")+
  ylab("Number of Wines")+
  scale_x_continuous(breaks = seq(1,10,1))+
  scale_y_continuous(breaks = seq(1,1000,50))+
  labs(title = "Distribution of Wine Quality:")
```

## Description One:

82.49% of wines in the dataset are of quality 5 and 6.  This graph is
important as we have to determine which other metrics lead to better wine
quality.

## Plot Two:

```{r echo=FALSE, message=FALSE, warning=FALSE, plot_two}
# Alcohol vs. Volatile Acidity vs. Quality.
ggplot(aes( y = alcohol, x = quality, color = volatile.acidity.partition),
  data = red_wine_ds)+
  geom_point()+
  geom_jitter()+
  geom_smooth(method = "lm", se = FALSE)+
  scale_x_continuous(breaks = seq(1, 10, 1))+
  scale_color_brewer("Volatile Acidity (g / dm^3)", palette = 'Blues')+
  xlab("Wine Quality")+
  ylab("Alcohol (% by Volume)")+
  labs(title = "Distribution of Alcohol % vs. Volatile Acidity with \
       Wine Quality:")+
  theme(panel.background = element_rect(fill = '#666666', colour = 'black'),
  panel.grid.major = element_line(colour = "black"),
  panel.grid.minor = element_line(colour = "black"),
  legend.key = element_rect(fill = '#666666'))
```

## Description Two:

Alcohol and volatile acidity are strongly correlated with the quality of the
wine.  We can observe that higher alcohol levels lead to good quality of wine.

## Plot Three:

```{r, echo=FALSE, message=FALSE, warning=FALSE, plot_three}
# Alcohol vs. Sulphates vs. Quality.
ggplot( aes(x = quality, y = alcohol, color = sulphates.partition),
  data = red_wine_ds)+
  geom_point()+
  geom_jitter()+
  geom_smooth(method = "lm", se = FALSE)+
  scale_color_brewer("Sulphates (g / dm3)", palette = 'Blues')+
  scale_x_continuous(breaks = seq(1, 10, 1))+
  xlab("Wine Quality")+
  ylab("alcohol (% by Volume)")+
  labs(title = "Distribution of Alcohol % vs. Sulphates with Wine Quality:")+
  theme(panel.background = element_rect(fill = '#666666', colour = 'black'),
  panel.grid.major = element_line(colour = "black"),
  panel.grid.minor = element_line(colour = "black"),
  legend.key = element_rect(fill = '#666666'))
```

## Description Three:

Alcohol and sulphates are positively correlated with the quality of the wine.
Higher values of sulphates and alcohol lead to good quality of wine.

------

<a name="reflection"></a>
**9. Reflection**

The aim of this analysis was to find out which chemical properties influence the
quality of red wines.  The dataset used had 1,599 observations and 12 variables.

The **Univariate Analysis** showed that 82.49% of the wines are of quality value
5 and 6. The other histograms in this section did not provide much help in
deciding what affected wine quality.

The **Bivariate Analysis** box plots of variables with respect to wine quality
showed that 4 variables - volatile acidity, alcohol, citric acid, and sulphates
- are involved in good quality wines.

The **Multivariate Analysis** scatter plots of variables correlated with wine
quality resulted in useful information. Good quality wine is produced with:

1. High sulphates and high alcohol.
2. High citric acid and high alcohol.
3. Low volatile acidity.
4. High sulphates.

*Future Analysis:*

The quality of the dataset is suspect.  82.49% of the wines are of quality value
5 and 6.  Ideally, a dataset where the quality variable was much closer to a
uniform distribution would be preferable.  A much larger number of observations
would also be preferable - tens of thousands or more, rather than the 1,599
found in the current dataset.
